BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-16 Msc-Slides/main.tex
Malware Analysis Preamble Removed due to Time Constraints
Slides re-worked according to Daniele's suggestions

\section{Introduction}
\subsection{Malware Analysis}

@@

    Two main types:
    \medskip
    \begin{itemize}
        \item \textbf{Static Analysis:}\\
        involves the inspection of the different data and code sections of a binary
        \item \textbf{Dynamic Analysis:}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{1cm}

@@

    \begin{textblock*}{2cm}(9.0cm,-6.5cm)
   \includegraphics[width=1.2cm]{image/search.png}% use the \includegraphics command here
	\end{textblock*}

@@

Functions can abstract
implementation details providing
a semantically richer representation of some functionality. \\\bigskip
Example:
    \begin{figure}
    	\vspace{-0.4cm}
        \includegraphics[scale=0.6]{image/fcmonitoring}
    \end{figure}

@@

A dynamic binary analysis
technique in which the behaviour of an application is inspected at run-time via the
injection of analysis code. 
\bigskip
\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}
\bigskip
\bigskip
\textbf{Problem 1}: existing products have limited logging capabilites

\subsection{Evasion}

    \begin{figure}
    	\vspace{-0.2cm}
    	\hspace*{1cm}
        \includegraphics[width=9cm]{image/evasive.pdf}
    \end{figure}

    \textbf{Problem 2}: API hooking techniques in literature are not coupled with mechanisms to hide their presence from evasive malware

@@

\begin{frame}
    \frametitle{Our solution: BlueTracer}

@@

\bigskip

Implementation details:
\begin{itemize}
\item Based on the \textbf{Intel Pin} DBI framework
\item Integrated with the \textbf{BluePill} stealthy execution framework
\item Combines reliable external sources of prototypes information
\end{itemize}

\bigskip

Key features:
\begin{itemize}
\item Undetected tracing of input parameters, output buffers and return values of over 17 000 system calls and library calls
\item Logging of asynchronous events
\item Resolution of named constants
\end{itemize}

@@

\subsection{Intel Pin}

@@

	Characteristics:
	\begin{itemize}
	\item \textbf{User-friendliness}
	\item \textbf{Portability}
	\item \textbf{Transparency}
	\item \textbf{Efficiency}
	\end{itemize}
	\bigskip
	\textbf{Analysis routines:} embody the code to be inserted during the application's execution\\
	\smallskip
	\textbf{Instrumentation routines:} determine where the analysis code has to be placed\\
	\bigskip
	Different analysis and instrumentation granularities
	\begin{itemize}
	\item Instruction, trace, routine and image
	\end{itemize}

@@

@@

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Evaluation with Al-Khaser}
	\textbf{Al-Khaser} is an open-source application which performs common checks employed by malware families to determine if they are being
executed in an analysis environment.
\\\medskip
Checks divided in categories:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Anti-Analysis}
\end{itemize}
\bigskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer was undetected with respect to all the checks!
\end{beamerboxesrounded}

    \begin{textblock*}{3cm}(7.5cm,-4.cm)
   \includegraphics[width=2cm]{image/lights.png}% use the \includegraphics command here
	\end{textblock*}

@@

File system artifacts can be checked in order to uncover the presence of a virtualized environment.
\vspace*{0.5cm}

@@

    \begin{table}[h]
\vspace*{0.1cm}
\begin{center}
\resizebox{9.2cm}{!}{
 \begin{tabular}{||c c c||}
 \hline
 \textbf{ID} & \textbf{MD5} & \textbf{Name} \\ 
 \hline\hline
 1 & 0af4ef5069f47a371a0caf22ae2006a6 & \textit{trojan/banker} \\ 
 \hline
 2 & 9437eabf2fe5d32101e3fbf9f6027880 & \textit{dropper} \\
 \hline
 3 & cbdda646a20d95f078393506ecdc0796 & \textit{trojan}\\
 \hline
 4 & cfdd16225e67471f5ef54cab9b3a5558 & Olympic\\
 \hline
 5 & ef694b89ad7addb9a16bb6f26f1efaf7 & CCleaner\\ 
 \hline
\end{tabular}
}
\end{center}
\end{table}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
The
logs collected by BlueTracer reveal behaviors consistent with the analysis reports
authored by Joe Security
\end{beamerboxesrounded}

@@

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and system call tracer for Windows programs specialized in evasive
malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Test with a larger set of highly evasive malware samples
	\item Improve logging capabilities
	\item Adopt log filtering techniques
	\end{itemize}

>>>>

\section{Introduction}

\subsection{Malware Analysis}

@@

    Two main types:
    \medskip
    \begin{itemize}
        \item \textcolor{sapienza}{\textbf{Static Analysis:}}\\
        involves the inspection of the different data and code sections of a binary
        \item \textcolor{sapienza}{\textbf{Dynamic Analysis:}}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{0.5cm}

@@

    \begin{textblock*}{2cm}(9.0cm,-6.1cm)
   \includegraphics[width=1.2cm]{image/search.png}% use the \includegraphics command here
	\end{textblock*}

    \begin{figure}
    	\vspace{-0.6cm}
    	\hspace*{0.5cm}
        \includegraphics[width=11cm]{image/evasive.pdf}
    \end{figure}

@@

Functions can abstract
implementation details providing
a semantically richer representation of some functionality. \\\bigskip
\textbf{Example}:
    \begin{figure}
    	\vspace{-0.4cm}
        \includegraphics[scale=0.6]{image/fcmonitoring}
    \end{figure}

@@

A dynamic binary analysis
technique in which the behaviour of an application is inspected at run-time, without the need of recompiling it, via the
injection of analysis code. 
\bigskip
\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}
\bigskip
\textcolor{sapienza}{\textbf{Problems:}}
\begin{enumerate}
\item Existing products have limited logging capabilites
\item API hooking techniques in literature are not coupled with mechanisms to hide their presence from evasive malware
\end{enumerate}

@@

\subsection{Overview}
\begin{frame}
    \frametitle{Our solution: BlueTracer}

\medskip

@@

\medskip

\textcolor{sapienza}{\textbf{Building blocks:}}
\begin{itemize}
\item Based on the \textbf{Intel Pin} DBI framework
\item Integrated with the \textbf{BluePill} stealthy execution framework
%\item Combines reliable external sources of prototypes information
\end{itemize}

\bigskip
	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=7cm]{image/BluePill.pdf}
    \end{figure}

@@

\begin{frame}
    \frametitle{BlueTracer: challenges and features}
\setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\textcolor{sapienza}{\textbf{Implementation challenges:}}

\begin{itemize}
\item Heterogeneity of Windows libraries used in malware and the lack of well-structured documentation for their prototypes
\begin{itemize}
\item[$\rightarrow$] Integration of reliable external sources \\ (\textbf{Dr. Memory}
and \textbf{CISCO PyREBox})
\end{itemize}
\item Making best use of
Intel Pin's capabilities with respect to run-time CPU and memory costs
\item Addressing some inherent limitations of Intel Pin

\end{itemize}

\textcolor{sapienza}{\textbf{Key features:}}
\begin{itemize}
\item Undetected tracing of input parameters, output buffers and return values of over 17 000 system calls and library calls
\item Logging of asynchronous events
\item Resolution of named constants
\end{itemize}

\iffalse
\subsection{Intel Pin}

@@

	Characteristics:
	\begin{itemize}
	\item \textbf{User-friendliness}
	\item \textbf{Portability}
	\item \textbf{Transparency}
	\item \textbf{Efficiency}
	\item \textbf{Robustness}
	\end{itemize}
	\medskip
	\textbf{Analysis routines:} embody the code to be inserted during the application's execution\\
	\smallskip
	\textbf{Instrumentation routines:} determine where the analysis code has to be placed\\
	\bigskip

%	Different analysis and instrumentation granularities
%	\begin{itemize}
%	\item Instruction, trace, routine and image
%	\end{itemize}

@@

\end{frame}
\fi

@@

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Evaluation with Al-Khaser}
	\textbf{Al-Khaser} is an open-source application which performs common checks employed by malware families to determine if they are being
executed in an analysis environment.
\\\medskip
Checks divided in categories:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Anti-Analysis}
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer: 
\begin{itemize}
\item remained undetected thanks to its integration with BluePill
\item managed to track all the checks
\end{itemize}
\end{beamerboxesrounded}

    \begin{textblock*}{3cm}(7.5cm,-5.5cm)
   \includegraphics[width=2cm]{image/lights.png}% use the \includegraphics command here
	\end{textblock*}

@@

\begin{beamerboxesrounded}[shadow=true]{}
Tracing a particular action of a malware instance allows to understand in detail what the sample's intentions are
\end{beamerboxesrounded}
\bigskip    
\textbf{Example}: file system artifacts can be checked in order to uncover the presence of a virtualized environment.
\vspace*{0.3cm}

@@

    \begin{table}[h]
\vspace*{-0.1cm}
\begin{center}
\resizebox{9.2cm}{!}{
 \begin{tabular}{||c c c||}
 \hline
 \textbf{ID} & \textbf{MD5} & \textbf{Name} \\ 
 \hline\hline
 1 & 0af4ef5069f47a371a0caf22ae2006a6 & \textit{trojan/banker} \\ 
 \hline
 2 & 9437eabf2fe5d32101e3fbf9f6027880 & \textit{dropper} \\
 \hline
 3 & cbdda646a20d95f078393506ecdc0796 & \textit{trojan}\\
 \hline
 4 & cfdd16225e67471f5ef54cab9b3a5558 & Olympic\\
 \hline
 5 & ef694b89ad7addb9a16bb6f26f1efaf7 & CCleaner\\ 
 \hline
\end{tabular}
}
\end{center}
\end{table}
\vspace{-0.1cm}
Evaluation was done manually and is a time-consuming process:
\begin{itemize}
\item Check if logs are congruous with Joe Security reports
\item Process Monitor as ground truth for system activity
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
The
logs collected by BlueTracer reveal behaviors consistent with the analysis reports
authored by Joe Security
\end{beamerboxesrounded}

@@

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and system call tracer for Windows programs specialized in evasive
malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Adoption of an automatic methodology for large-scale evaluation 
	\item Improvement of logging capabilities
	\item Usage of log filtering techniques
	\end{itemize}

### 2018-07-16 conclusions.tex
Slides re-worked according to Daniele's suggestions

The obvious next step in BlueTracer's development is to test it with larger a set of highly evasive malware samples, in order to further evaluate its effectiveness in a real-life scenario.

>>>>

The obvious next step in BlueTracer's development would be to introduce an automatic methodology for large-scale evaluation so that the tool can be tested with a larger a set of highly evasive malware samples, in order to further evaluate its effectiveness in a real-life scenario.
